---
title: "jupyter_test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(patchwork)
library(extrafont)


```


get the daily data
```{r}


url <- "http://www.vdh.virginia.gov/content/uploads/sites/182/2020/03/VDH-COVID-19-PublicUseDataset-Cases.csv"
dest <- paste(getwd(), "/data/va_daily.csv",sep="")

download.file(url, dest)


```

read data and select/rename desired fields
```{r}
data <- read_csv("data/va_daily.csv" ,
                col_types = cols('Report Date' = col_date(format = "%m/%d/%Y"),
                                 'Total Cases' = col_integer(),
                                 'Hospitalizations' = col_integer(),
                                 'Deaths' = col_integer()))
                

data <- data %>% select(date=`Report Date`, Locality, district=`VDH Health District`, cases=`Total Cases`, Hospitalizations, Deaths)

data
```


Filter for Fairfax Health District then calculate new cases, new hospitalizations, and new deaths.

Add fields to record sign of new cases differences and then add a column to calculate the streak of days with decreasing or
zero new cases.
```{r}

# Uncomment and use this code if graphing Fairfax Health District 
# fairfax <- data %>% filter(district == 'Fairfax') %>% 
#   #group_by(date, district) %>%
#   #group_by(date) %>% 
#   summarize(tot_cases = sum(cases), hosp=sum(Hospitalizations),dead=sum(Deaths)) %>% 
#   arrange(date) %>% 
#   #ungroup() %>% 
#   #group_by(district) %>%
#   select(date, Locality, tot_cases, hosp, dead) %>%
#   mutate(new_cases = tot_cases - lag(tot_cases,1),
#          new_hosp = hosp - lag(hosp,1),
#          new_dead = dead - lag(dead,1)) %>% 
#   # 1 means 0 or negative difference,  0 means positive difference
#   mutate(d_cases = if_else(sign(new_cases - lag(new_cases,1)) ==1,0,1,missing=0)) %>%
#   group_by(grp=cumsum(d_cases==0)) %>% 
#   mutate(streak = ifelse(d_cases==1, lag(cumsum(d_cases==1))+1,0))


#Use this code if graphing Fairfax County (not F.Church or Fairfax City) 
fairfax <- data %>% filter(Locality == 'Fairfax') %>% 
  select(date,cases,hosp=Hospitalizations,dead=Deaths ) %>% 
  arrange(date) %>%
  mutate(new_cases = cases - lag(cases,1),
         new_hosp = hosp - lag(hosp,1),
         new_dead = dead - lag(dead,1)) %>%
  # 1 means 0 or negative difference,  0 means positive difference, d_cases stands for difference in new cases
  mutate(d_cases = if_else(sign(new_cases - lag(new_cases,1)) ==1,0,1,missing=0)) %>%
  group_by(grp=cumsum(d_cases==0)) %>%
  mutate(streak = ifelse(d_cases==1, lag(cumsum(d_cases==1))+1,0))


fairfax

```

Read in the trend line from the stan model

```{r}
stan <- read_csv("trends/trend_today.csv")

```



Plot everything
```{r fig.height=10, warning=FALSE}

#grab the maximum date
today <- max(fairfax$date)

# plot new cases
n <- ggplot(fairfax, aes(x=date, y=new_cases)) +
  geom_bar(stat="identity",fill="grey49", alpha=0.75) +
  geom_line(data=stan, aes(x=date, y=trend), color="firebrick", size=2, linetype="solid")+
  geom_text(data=subset(fairfax, date ==max(date)), aes(y=new_cases, label=new_cases),
            position = position_dodge(width = 0.5), vjust=-.5, color="grey32") +
  theme_minimal()+
  theme(text = element_text(family='Verdana'),
        plot.title = element_text(size = 12, face="bold"),
        panel.grid.major = element_line(color="grey68")) +
  labs(title="New Covid-19 Cases",
       subtitle = sprintf("Total Cases: %s", max(fairfax$cases)),
           x=NULL, y=NULL) +
  xlim(date("2020-03-17"), max(fairfax$date)+1)



# plot case streak
a <- ggplot(fairfax, aes(x=date, y=streak)) + 
  geom_bar(stat="identity", fill="red",alpha=0.75) +
  geom_text(data=subset(fairfax, date ==max(date)), aes(y=streak, label=streak),
            position = position_dodge(width = 0.5), vjust=-.5, color="grey32") +
  ylim(0,15) +
  theme_minimal()+
  theme(text = element_text(family='Verdana'),
        plot.title = element_text(size = 12, face="bold"),
        panel.grid.major = element_line(color="grey68")) +
  labs(title="Consecutive Days with Fewer New Covid-19 Cases",
       x=NULL, y=NULL) +
  xlim(date("2020-03-17"), max(fairfax$date)+1)

# plot hospitalizations
h <-  ggplot(fairfax, aes(x=date, y=new_hosp)) +
  geom_bar(stat="identity", fill = "purple", alpha=0.75, na.rm=TRUE) +
  geom_text(data=subset(fairfax, date ==max(date)), aes(y=new_hosp, label=new_hosp),
            position = position_dodge(width = 0.5), vjust=-.5, color="grey32") +
  theme_minimal()+
  theme(text = element_text(family='Verdana'),
        plot.title = element_text(size = 12, face="bold"),
        panel.grid.major = element_line(color="grey68")) +
  labs(title="New Covid-19 Hospitalizations",
       subtitle = sprintf("Total Hospitalizations: %s", format(fairfax$hosp[fairfax$date==today],big.mark=",", scientific=FALSE)),
       x=NULL, y=NULL) +
  ylim(0, (max(fairfax$new_hosp)+10)) +
  xlim(date("2020-03-17"), max(fairfax$date)+1)


# plot deaths
d <- ggplot(fairfax, aes(x=date, y=new_dead)) +
  geom_bar(stat="identity", fill = "brown",alpha=0.75,  na.rm=TRUE) +
  geom_text(data=subset(fairfax, date ==max(date)), aes(y=new_dead, label=new_dead),
            position = position_dodge(width = 0.5), vjust=-.5, color="grey32") +
  theme_minimal()+
  theme(text = element_text(family='Verdana'),
        plot.title = element_text(size = 12, face="bold"),
        panel.grid.major = element_line(color="grey68")) +
  labs(title="New Covid-19 Deaths",
       subtitle = sprintf("Total Deaths: %s", fairfax$dead[fairfax$date==today]),
               x=NULL, y=NULL) +
  ylim(0, (max(fairfax$new_dead)+10)) +
  xlim(date("2020-03-17"), max(fairfax$date)+1)



# arrange the plots vertically
patchwork <- n / a / h/ d 

#create final plot
# final <- patchwork + 
#   plot_layout(heights = c(2,1,1,1))+
#   plot_annotation(title=sprintf("Fairfax Health District: Covid-19 Summary for %s %s %s",
#                                 day(today), month.name[month(today)],year(today)),
#                   subtitle= "Fairfax Health District includes Fairfax County, City of Fairfax, and City of Falls Church\nSource: Virginia Department of Health",
#                   theme = theme(plot.title = element_text(size = 16))) &
#   theme(text = element_text(family='Verdana'),
#         plot.caption = element_text(face = 'italic')) 

final <- patchwork + 
  plot_layout(heights = c(2,1,1,1))+
  plot_annotation(title=sprintf("Fairfax County: Covid-19 Summary for %s %s %s",
                                day(today), month.name[month(today)],year(today)),
                  subtitle= "Charts exclude numbers for City of Fairfax and City of Falls Church\nSource: Virginia Department of Health",
                  theme = theme(plot.title = element_text(size = 16))) &
  theme(text = element_text(family='Verdana'),
        plot.caption = element_text(face = 'italic')) 

final
```

```{r}
ggsave("covid_daily_ffx.png",width = 8.5, height=11.0)
```





